---
title: "Hotel Booking Analysis"
format: html
---


```{r packages}
library(tidyverse)
library(caret)
library(googledrive)
library(GGally)
library(infer)
library(ggthemes)
library(rsample)
library(ggfortify)
library(ggmosaic)
theme_set(theme_tufte())
```


```{r data}
hotel = read_csv("Hotel Reservations.csv")
```



```{r}
hotel  |> 
    skimr::skim()
```


It appears as though there are not any missing values which is good, the next question is exploring the distributions of these variables.

We should probably start with our supervisor here, which I believe is booking status.

```{r}
hotel  |>  
    ggplot(aes(x = booking_status)) +
    geom_bar()
```

It appears as though the majority were not canceled, though the classes are fairly balanced.

Let's explore all of the categorical columns real quick

```{r}
hotel  |> 
    keep(is.character)   |> 
    select(-Booking_ID)  |> 
    gather()  |> 
    ggplot(aes(x = value)) +
    facet_wrap(~key, scales = "free") +
    geom_bar() + 
    coord_flip()
```

Now the numeric columns

```{r}
hotel  |> 
    keep(is.numeric)   |> 
    gather()  |> 
    ggplot(aes(x = value)) +
    facet_wrap(~key, scales = "free") +
    geom_histogram() +
    coord_flip()
```

Let's go ahead and add the Booking status to study potential relationships with the features

```{r}
hotel  |> 
    keep(is.character)   |> 
    select(-Booking_ID)  |> 
    pivot_longer(cols = type_of_meal_plan:market_segment_type,
    names_to = "columns")  |> 
    ggplot(aes(y = value, fill = booking_status)) +
    facet_wrap(~columns, scales = "free") +
    geom_bar(position = "fill")
```

There does appear to be some difference in proportions based off of the categorical. Complementary never seems to cancel whereas Onlien and Meal plan 2 seem to have the highest incidence of canceling.

```{r}
hotel  |> 
    select(where(is.numeric), booking_status)  |> 
    pivot_longer(cols = no_of_adults:no_of_special_requests,
    names_to = "columns")  |> 
    ggplot(aes(x = value, fill = booking_status)) +
    facet_wrap(~columns, scales = "free", nrow = 2) +
    geom_histogram(position = 'dodge')
```

It appears as though there really isn't too much to differentiate them in terms continuous.

```{r}
hotel  |> 
    select(where(is.numeric), booking_status)  |> 
    pivot_longer(cols = no_of_adults:no_of_special_requests,
    names_to = "columns") |> 
    group_by(columns, booking_status) |> 
    summarise(group_means = mean(value)) |> 
  pivot_wider(names_from = booking_status, values_from = group_means)
```

I want to do a t-test on each of these groups to help understand if there are significant differences between the different groups

```{r}
hotel |> 
  t_test(formula = avg_price_per_room ~ booking_status,
         alternative = "two-sided") |> 
  mutate(variable = "avg_price_per_room")
```

Let's go ahead and prototype this using a string

```{r t_test_proto}
var_name = "avg_price_per_room"
response = "booking_status"
hotel |> 
  t_test(formula = paste(var_name, " ~ ", response) |> as.formula(),
         alternative = "two-sided") |> 
  mutate(variable = var_name)
```
We now create the function

```{r t_test_func}
auto_t_test <- function(data, var_name, factor, order_vec){
  df_out <- data |> 
    t_test(formula = paste(var_name, " ~ ", factor) |> as.formula(),
          order = order_vec,
         alternative = "two-sided") |>
  mutate(variable = var_name)
  return(df_out)
}

```

We now attempt to apply this to all of the numeric columns

```{r}
numeric_cols <- hotel |> 
  keep(is.numeric) |> 
  colnames()
t_hotel <- sapply(numeric_cols, FUN = auto_t_test, data = hotel, factor = "booking_status", 
       order_vec =  c("Canceled", "Not_Canceled")) |> 
  t() |> 
  apply(MARGIN= 2, FUN = unlist) |> 
  
  as_tibble()  |> 
  select(-alternative) |> 
  relocate(variable) |> 
  mutate(across(.cols = statistic:upper_ci, ~ as.numeric(.x)))
```

Getting all groups that are greater

```{r}
t_hotel |> 
  filter(p_value < .05 / n() & estimate > 0) |> 
  select(variable, estimate) |> 
  arrange(desc(estimate))
```

```{r}
t_hotel |> 
  filter(p_value < .05 / n() & estimate < 0) |> 
  select(variable, estimate) |> 
  arrange(desc(estimate))
```
Let's investigate the lead_time a little bit more

```{r}
hotel |> 
  ggplot(aes(x = lead_time, y = booking_status, fill = booking_status)) +
  geom_violin()
```

Now onto the categorical variables.

```{r cat_proto}
hotel |> 
  ggplot() +
  geom_mosaic(aes(x = product(type_of_meal_plan), fill = booking_status)) +
  theme_mosaic() +
  scale_fill_manual(values = c( "#ABD9E9", "#FEE090")) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  
```


```{r}
hotel |> 
  ggplot() +
  geom_mosaic(aes(x = product(room_type_reserved), fill = booking_status)) +
  theme_mosaic() +
  scale_fill_manual(values = c( "#ABD9E9", "#FEE090")) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```




```{r}
hotel |> 
  ggplot() +
  geom_mosaic(aes(x = product(market_segment_type), fill = booking_status)) +
  theme_mosaic() +
  scale_fill_manual(values = c( "#ABD9E9", "#FEE090")) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Chi-Squared Test for categorical

```{r}
# calculate the observed statistic
observed_indep_statistic <- hotel %>%
  specify(booking_status ~ day_of_week) %>%
  hypothesize(null = "independence") %>%
  calculate(stat = "Chisq")
observed_indep_statistic
```


```{r}
hotel_clean %>%
  specify(booking_status ~ day_of_week) %>%
  assume(distribution = "Chisq") %>%
  visualize() + 
  shade_p_value(observed_indep_statistic,
                direction = "greater")
```


Lastly, let's explore what to do exactly about these data variables

```{r}
hotel |> 
  ggplot(aes(x = arrival_month, color = booking_status)) +
  geom_line(stat = "count", linewidth = 2)
```

```{r}
hotel |> 
  group_by(arrival_month) |> 
  count(booking_status) |> 
  ungroup() |> 
  pivot_wider(names_from = booking_status, values_from = n) |> 
  mutate(total = Canceled + Not_Canceled,
        prop_canceled = Canceled / (total)) |> 
  ggplot(aes(x = arrival_month, y = prop_canceled)) +
  geom_point(aes(size = total), color = "maroon") +
  geom_line()
```


```{r}
hotel |> 
  group_by(arrival_year) |> 
  count(booking_status) |> 
  ungroup() |> 
  pivot_wider(names_from = booking_status, values_from = n) |> 
  mutate(total = Canceled + Not_Canceled,
        prop_canceled = Canceled / (total)) |> 
  ggplot(aes(x = arrival_year, y = prop_canceled)) +
  geom_point(aes(size = total), color = "maroon") +
  geom_line()
```

This could probably be an indicator variable.

```{r}
hotel |>
  group_by(arrival_date) |> 
  count(booking_status) |> 
  ungroup() |> 
  pivot_wider(names_from = booking_status, values_from = n) |> 
  mutate(total = Canceled + Not_Canceled,
        prop_canceled = Canceled / (total)) |> 
  ggplot(aes(x = arrival_date, y = prop_canceled)) +
  geom_point(aes(size = total), color = "maroon") +
  geom_line()

```

Day of the week is probably gonna be helpful here

```{r}
hotel <- hotel |> 
  mutate(arrival_dt = ymd(paste(arrival_year, arrival_month, arrival_date ,sep="-"))) |> 
  filter(!is.na(arrival_dt))
```


```{r}
hotel <- hotel |> 
  mutate(day_of_week = weekdays(arrival_dt))
```


```{r}
hotel |> 
  ggplot() +
  geom_mosaic(aes(x = product(day_of_week), fill = booking_status))
```

```{r}
hydroTSM::time2season(hotel$arrival_dt, out.fmt = "seasons")
hotel <- hotel |> 
  mutate(season = hydroTSM::time2season(arrival_dt, out.fmt = "seasons"))
```


```{r}
hotel |> 
  ggplot() +
  geom_mosaic(aes(x = product(season), fill = booking_status))
```

```{r}
hotel |> 
  mutate(day_type = if_else(day_of_week %in% c("Saturday", "Sunday"), "Weekend", "Weekday")) |> 
  ggplot() +
  geom_mosaic(aes(x = product(day_type), fill = booking_status))
```

I'm not sure if type of day is that much more useful than just the day.

So, season, year and then day type are useful and more actionable as opposed to the raw numbers.

```{r data_processing}
hotel_clean <- hotel |> 
  filter(arrival_date != 29 & arrival_month!= 2) |> 
  mutate(arrival_dt = ymd(paste(arrival_year, arrival_month, arrival_date ,sep="-")),
         day_of_week = weekdays(arrival_dt),
         day_type = if_else(day_of_week %in% c("Saturday", "Sunday"), "Weekend", "Weekday"),
         total_people = no_of_children + no_of_adults,
         season = hydroTSM::time2season(arrival_dt, out.fmt = "seasons"),
         across(where(is.character), as.factor))
hotel_clean
```


```{r}
write_rds(x = hotel_clean, file = "Data/hotel.rds")
```


```{r}
hotel_clean |> 
  ggplot() +
  ggforce::geom_autopoint(aes(type_of_meal_plan, room_type_reserved))
```

# Works pretty well for discrete versus discrete

```{r}
# Discrete and continuos
hotel_clean |> 
  ggplot() +
  ggforce::geom_autopoint(aes(type_of_meal_plan, avg_price_per_room))
```

# Continuos versus Continous

```{r}
hotel_clean |> 
  ggplot(aes(lead_time, avg_price_per_room)) +
  ggforce::geom_autopoint(aes(lead_time, avg_price_per_room), color = "maroon", position = "jitter") +
  xlab("Time ") +
  ylab("Average Price per Room")
```

```{r}
ggpairs(data = hotel_clean,  columns = 3:5, aes(color = day_type), columnLabels = c("# Children", "# Adults", "# Weeknights"), lower = list(continuous = ggally_count, combo = ggally_dot_no_facet)) +
  theme_base()
```

```{r}
list.files("R", full.names = TRUE) %>% walk(source)
```


```{r}
hotel_clean |> 
  ggplot_two_columns(col1 = no_of_adults, col2 = no_of_children)
```

```{r}

```

